Shader Pipeline and Effect Encapsulation using CLOS

### 2019-02-27 els2019-shader-pipeline/paper.tex
Incorporate some peer review feedback

\begin{abstract}
  Modern real-time graphics make use of a lot of tricks in order to produce stunning visuals. Many of these tricks require separate rendering passes, as well as separate rendering logic for each pass. These passes are then combined in a variety of ways in order to produce a final image. The interaction between such a rendering pass and the objects it draws, as well as the interaction between multiple passes within a pipeline can become quite complex. Often times, in order to handle this complexity, the passes and objects are generalised, and the render logic is left to be executed almost entirely with either the object or the pass. We present a new method of representing objects, passes, and pipelines, which allows a modular encapsulation of effects and rendering behaviour, as well as object-oriented composition through inheritance. We make use of CLOS' multiple inheritance, multimethods, and standard method combination to form extensible protocols that allow this new method. We also make use of the MOP, in order to introduce additional metadata to classes.
\end{abstract}

@@

In order to permit this amount of control, we make use of the shader composition capabilities that we presented in our previous work\cite{glsloop}. This technique allows us to combine pieces of GLSL code. On top of this shader combination sits a new protocol to control the interaction between shader passes and objects. \\

@@

This protocol thus allows a great amount of control both over the effective shader code used to render an object, as well as the behaviour leading up towards the actual draw call for an object.

@@

\href{https://raw.githubusercontent.com/Shinmera/talks/master/els2019-shader-pipeline/pipeline-allocation.gif}{https://raw.githubusercontent.com/Shinmera/talks/master/\\els2019-shader-pipeline/pipeline-allocation.gif}

@@

>>>>

\begin{abstract}
  Modern real-time graphics make use of a lot of tricks in order to produce stunning visuals. Many of these tricks require separate rendering passes, as well as separate rendering logic for each pass. These passes are then combined in a variety of ways in order to produce a final image. The interaction between such a rendering pass and the objects it draws, as well as the interaction between multiple passes within a pipeline can become quite complex. Often times, in order to handle this complexity, the passes and objects are generalised, and the render logic is controlled almost entirely by either the object or the pass. We present a new method of representing objects, passes, and pipelines, which allows a modular encapsulation of effects and rendering behaviour, as well as object-oriented composition through inheritance. We make use of CLOS' multiple inheritance, multimethods, and standard method combination to form extensible protocols that allow this new method. We also make use of the MOP, in order to introduce additional metadata to classes.
\end{abstract}

@@

In order to permit this amount of control, we make use of the shader composition capabilities that we presented in our previous work\cite{glsloop}. This technique allows us to combine pieces of OpenGL Shader Language (GLSL) code. On top of this shader combination sits a new protocol to control the interaction between shader passes and objects. \\

@@

This protocol thus allows a great amount of control both over the effective shader code used to render an object, as well as the behaviour leading up towards the actual draw call for an object. \\

The overall protocol thus only invokes overhead for the dispatch of the \code{render-with}, \code{render}, and \code{shader-program-for-pass} generic functions for each rendered object per frame. All other functions are only invoked during loading, or in exceptional situations. In our use of the system so far we have not found this to impact performance significantly.

@@

\href{https://raw.githubusercontent.com/Shinmera/talks/master/els2019-shader-pipeline/pipeline-allocation.gif}{https://raw.githubusercontent.com/Shinmera/talks/master/\\els2019-shader-pipeline/pipeline-allocation.gif} \\

With this algorithm we reduce the number of necessary textures to allocate down to five. With a primitive allocation of one texture per output, we would instead end up with seven textures. Note that the shadow map output is not assigned, as it requires a different type of texture from the rest -- a depth texture. This texture would be allocated in a second pass.

@@

\newpage ~ % Fucking latex, give me a new page.
